Fast Radio Burst Cosmology

### 2014-04-29 writeup/dm_angular_ps.tex
Fixed up some equation, clearified set notation.

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\{D_i\}$ with known angular positions but with unknown
redshifts depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\{z_i\})$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\{z_i\}) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $\{D^{\bar{n_e}}_i\}$ are
uncorrelated:
\begin{align}
\langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
    - \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle
    =& \delta_{ij} \int_0^\infty \ud z p(z) \int_0^{\eta(z)} \ud \eta'
        (1 + z'(\eta')) \textrm{(not quite right, subtract off the mean.)}
        \equiv& \sigma_{\bar{n}_e}^2
\end{align}

We further assume that $\{D^{\rm src}_i\}$ are uncorrelated. Again this may not
be strictly true due to clustering of FRB sources and again this may bias our
estimator. We parameterized the variance from this term as 
\begin{equation}
    \langle D^{\rm src}_i D^{\rm src}_j \rangle
    - \langle D^{\rm src}_i \rangle \langle D^{\rm src}_j \rangle
    = \delta_{ij} \sigma^2_{\rm src}.
\end{equation}

The cosmological information lies in the correlations within the set
$\{D^{\delta_e}_i\}$. If the redshifts of the observed FRBs where known, we
would write the probability density function of this set as
$p(\{D^{\delta_e}_i\} | \{z_i\})$, which has second moments
\begin{equation}
    w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j) \equiv \int \ud \{D^{\delta_e}_i\}
        D^{\delta_e}_i D^{\delta_e}_j p(\{D^{\delta_e}_i\} | \{z_i\}).
\end{equation}
$w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j)$ is the redshift dependant DM angular
correlation function, whose dependence on cosmology is derived below.
Marginalizing $p(\{D^{\delta_e}_i\} | \{z_i\})$ over the unknown redshifts
yields
\begin{equation}
    p(\{D^{\delta_e}_i\}) = \int_0^\infty \ud \{z_i\}
        p(\{z_i\}) p(\{D^{\delta_e}_i\} | \{z_i\}).
\end{equation}
The second moment of the marginalized distribution is
\begin{align}
    \langle D^{\delta_e}_i D^{\delta_e}_j \rangle =& \\
        =& \textrm{stuff} \\
    =& \int \ud z \ud z' p(z) p(z') w^{z z'}(\hat{n}_i\cdot\hat{n}_j) \qquad (i \neq j) \\
    =& \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(\hat{n}_i\cdot\hat{n}_j)
    \qquad (i = j)
\end{align}
I think.

The covariance matrix is
\begin{align}
    C_{ij} \equiv& \langle D_i D_j \rangle - \langle D_i \rangle \langle D_j \rangle
           =& \textrm{a bunch of things we know (special case for $i=j$)}.
           =& w(\hat{n}_i\cdot\hat{n}_j) \quad (i \neq j)
    =& \sigma_{\bar{n}_e}^2 + \sigma^2_{\rm src} + \textrm(the $w$ term) \quad (i = j)
\end{align}

Now that we have the covariance matrix, we construct a quadratic estimator for
a chosen set of parameters $\{\lambda_\alpha\}$. The estimator is Equation~11.94 from
Dodelson, although we have to modify it slightly to exclude the
auto-correlations.  The error (assuming Gaussianity) is Equation~11.100.

>>>>

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\boldsymbol D$ (whose members are $D_i$)
with known angular positions, $\hat n_i$, but with unknown
redshifts, $z_i$, depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\boldsymbol z)$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\boldsymbol z) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $D^{\bar{n}_e}_i$ have
uncorrelated probability distributions
\begin{align}
    p(D^{\bar{n}_e}_i)
    &= p(z_i) \left|\frac{\ud D^{\bar{n}_e}}{\ud z} \right|_{z_i}^{-1} \\
        &= p(z_i) \left(\frac{\ud \eta}{\ud z} \right)_{z_i}^{-1},
\end{align}
whose second moment is
\begin{align}
    \langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
        &- \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle \\
        &= \delta_{ij} \left[ \int_0^\infty \ud D^{\bar{n}_e} (D^{\bar{n}_e})^2
            p(D^{\bar{n}_e})
            - \left( \int_0^\infty \ud D^{\bar{n}_e} D^{\bar{n}_e}
        p(D^{\bar{n}_e}) \right)^2 \right]\\
        &\equiv \delta_{ij} \sigma_{\bar{n}_e}^2\\
\end{align}

We further assume that $D^{\rm src}_i$ are uncorrelated. Again this may not
be strictly true due to clustering of FRB sources and again this may bias our
estimator. We parameterized the variance from this term as 
\begin{equation}
    \langle D^{\rm src}_i D^{\rm src}_j \rangle
    - \langle D^{\rm src}_i \rangle \langle D^{\rm src}_j \rangle
    = \delta_{ij} \sigma^2_{\rm src}.
\end{equation}

The cosmological information lies in the correlations between the
$D^{\delta_e}_i$. If the redshifts of the observed FRBs where known, we
would write the probability density function of this set as
$p(\boldsymbol D^{\delta_e} | \boldsymbol z)$, which has second moments
\begin{equation}
    w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j) \equiv \int \ud \boldsymbol D^{\delta_e}
        D^{\delta_e}_i D^{\delta_e}_j p(\boldsymbol D^{\delta_e} | \boldsymbol z).
\end{equation}
$w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j)$ is the redshift dependant DM angular
correlation function, whose dependence on cosmology is derived below.
Marginalizing $p(\boldsymbol D^{\delta_e} | \boldsymbol z)$ over the unknown redshifts
yields
\begin{equation}
    p(\boldsymbol D^{\delta_e}) = \int_0^\infty \ud \boldsymbol z
        p(\boldsymbol z) p(\boldsymbol D^{\delta_e} | \boldsymbol z).
\end{equation}
The second moment of the marginalized distribution is
\begin{align}
    \langle D^{\delta_e}_i D^{\delta_e}_j \rangle =& \\
        =& \textrm{stuff} \\
    =& \int \ud z \ud z' p(z) p(z') w^{z z'}(\hat{n}_i\cdot\hat{n}_j) \qquad (i \neq j) \\
    =& \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(\hat{n}_i\cdot\hat{n}_j)
    \qquad (i = j)
\end{align}
I think.

The covariance matrix is
\begin{align}
    C_{ij} \equiv& \langle D_i D_j \rangle - \langle D_i \rangle \langle D_j
    \rangle\\
           =& w(\hat{n}_i\cdot\hat{n}_j) \quad (i \neq j)\\
    =& \sigma_{\bar{n}_e}^2 + \sigma^2_{\rm src} + \textrm{(the $w$ term)} \quad (i = j)
\end{align}

Now that we have the covariance matrix, we construct a quadratic estimator for
chosen parameters $\lambda_\alpha$. The estimator is Equation~11.94 from
Dodelson, although we have to modify it slightly to exclude the
auto-correlations.  The error (assuming Gaussianity) is Equation~11.100.
